Diversity dynamics in Nymphalidae butterflies: Effect of phylogenetic uncertainty on diversification rate shift estimates

### 2015-01-14 response_to_reviewers.tex

\textbf{21. Reviewer's Comment:}
Pages 7-8, fourth paragraph of Discussion: Per previous comment of BAMM analysis, this paragraph should probably be removed. If it is not, both shifts detected by BAMM analysis (not just one) should be described and compared to the four identified in MultiMEDUSA (are they at the same nodes or different nodes?).

\textbf{21. Response:}

\begin{quote}
\color{blue}
We removed the paragraph according to the reviewer's suggestion.
\end{quote}

\textbf{22. Reviewer's Comment:}
Page 8, Hostplant use and diversification... section, Ithomiini subsection,
fifth paragraph, first sentence: insert 'rate': “The increase in diversification
rate inferred by MEDUSA...”

\textbf{22. Response:}

\begin{quote}
\color{blue}
We included the word ``rate'' according to the reviewer's suggestion.
\end{quote}

\textbf{23. Reviewer's Comment:}
Page 8, Hostplant use and diversification... section, Ithomiini subsection,
sixth paragraph, second sentence: replace “might not be too efficient” with 
“may be inefficient”.

\textbf{23. Response:}

\begin{quote}
\color{blue}
We reworded the sentence according to the reviewer's suggestion.
\end{quote}

### 2015-01-14 MS_revised.tex
comment 23

Although the Solanaceae genera used by the Ithomiini clades are well
known \cite{willmott2006}, we do not have any understanding on the physiological
routes involved in the detoxification of Solanaceae compounds by the
several lineages of Ithomiini. We can speculate that older lineages
exploiting a novel toxic resource \cite{willmott2006, wahlberg2009} might not be too efficient
in metabolizing plant toxins and that younger lineages are able to deal
with toxins more efficiently, so that host switching events within
Solanaceae are possible, which can lead to higher diversification.
Studies in \emph{Papilio} species have reported that detoxification
enzymes can become more efficient in metabolizing toxins than ancestral
configurations of the proteins, providing more opportunities for
hostplant switch \cite{li2003}. This might be the reason why the basal
Ithomiini subtribes Melinaeina and Mechanitina are so species-poor and
restricted to few Solanaceae hosts \cite{willmott2006}, while recent subtribes are
species-rich and have expanded their host range into several Solanaceae
lineages \cite{willmott2006}. It might be that the switch to feeding in Solanaceae
was an important event in the evolutionary history of Ithomiini, but the
actual radiation occurred after critical physiological changes (a
probable key innovation) allowed efficient detoxification of Solanaceae
toxins.

>>>>

Although the Solanaceae genera used by the Ithomiini clades are well
known \cite{willmott2006}, we do not have any understanding on the physiological
routes involved in the detoxification of Solanaceae compounds by the
several lineages of Ithomiini. We can speculate that older lineages
exploiting a novel toxic resource \cite{willmott2006, wahlberg2009} 
may be inefficient
in metabolizing plant toxins and that younger lineages are able to deal
with toxins more efficiently, so that host switching events within
Solanaceae are possible, which can lead to higher diversification.
Studies in \emph{Papilio} species have reported that detoxification
enzymes can become more efficient in metabolizing toxins than ancestral
configurations of the proteins, providing more opportunities for
hostplant switch \cite{li2003}. This might be the reason why the basal
Ithomiini subtribes Melinaeina and Mechanitina are so species-poor and
restricted to few Solanaceae hosts \cite{willmott2006}, while recent subtribes are
species-rich and have expanded their host range into several Solanaceae
lineages \cite{willmott2006}. It might be that the switch to feeding in Solanaceae
was an important event in the evolutionary history of Ithomiini, but the
actual radiation occurred after critical physiological changes (a
probable key innovation) allowed efficient detoxification of Solanaceae
toxins.
